### 2021-07-21 paper/paper.tex
WIP: overview of Trixi.jl

% Physicists rule
\usepackage{siunitx}
% \sisetup{output-exponent-marker=\text{e}}
% \sisetup{tight-spacing=true}

@@

% Macros
\newcommand{\trixi}{Trixi.jl}

@@

\begin{itemize}
  \item Basic design of \trixi and its capabilities
  \item PID comparison of Fluxo and \trixi
  \item Discussion: Julia for numerical simulation science
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
\end{itemize}

@@

@@

Finally, the different mesh types available in \trixi allow further optimizations.
For example, if the computational domain is essentially a cube (or square, line)
without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
be used. Depending on the relative computational effort of the discretization
(such as the polynomial degree and the choice of the volume terms), the
Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
more efficient than the curvilinear meshes in \trixi.

@@

TODO %TODO
\begin{itemize}
  \item Some nice features
  \item Compilation time can be annoying but also allows cool features and performance
  \item Extendability: Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}

>>>>

% Mathematics is necessary
\usepackage{amsmath}
\allowdisplaybreaks

% Physics rules
\usepackage{siunitx}
% \sisetup{output-exponent-marker=\text{e}}
% \sisetup{tight-spacing=true}

@@

% Macros
\usepackage{xspace}
\newcommand{\ca}[0]{{ca.\@}\xspace}
\newcommand{\cf}[0]{{cf.\@}\xspace}
\newcommand{\eg}[0]{{e.g.\@}\xspace}
\newcommand{\ie}[0]{{i.e.\@}\xspace}
\newcommand{\etc}[0]{{etc.\@}\xspace}
\newcommand{\trixi}{Trixi.jl\xspace}

@@

\begin{itemize}
  \item Basic design of \trixi and its capabilities
  \item PID comparison of Fluxo and \trixi
  \item Discussion: Julia for numerical simulation science
  \item Prepare repro repo including scripts for \trixi and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
  \item The title of our talk at JuliaCon is
  ``Adaptive and extendable numerical simulations with Trixi.jl''.
  Shall we use something like
  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
  A case study of Julia for scientific computing'' for this paper?
  \item Restructure non-conservative stuff and add MHD benchmarks?
  \item Shall we re-create the schematic overview in tikz?
\end{itemize}

@@

\trixi is designed as a library of high-order methods for hyperbolic PDEs of
the form
\begin{equation}
\label{eq:hcl}
  \partial_t u(t, x) + \sum_{j=1}^d \partial_{x_j} f^j(u) = s(t, x, u),
  \quad t \in (0, T), x \in \Omega,
\end{equation}
in $d \in \{1, 2, 3\}$ space dimensions. Here, the independent variables are
time $t$ and space coordinates $x \in \Omega \subset \mathbb{R}^d$. The conserved
quantities are denoted as $u$, \eg, mass, momentum, and energy for the compressible
Euler equations of an ideal gas.
The physical model is specified by the fluxes $f^j$ and the source term $s$.
In addition, suitable initial and boundary conditions (ICs, BCs) are required.
\trixi also handles non-conservative source terms containing derivatives of the
conserved quantities as in the shallow water equations or magnetohydrodynamics
with divergence cleaning.

\subsection{High-level overview of the structure}

\trixi is based on the method of lines. Thus, a discretization of \eqref{eq:hcl}
is obtained in two steps. First, a spatial semidiscretization is created. Next,
the resulting ordinary differential equation (ODE) is solved using a time
integration method. Currently, \trixi focuses on the spatial semidiscretization
and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, a part
of DifferentialEquations.jl \cite{rackauckas2017differentialequations}.

\begin{figure*}
  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
  \caption{Schematic overview of the basic components of \trixi and how they
           interact with each other.}
  \label{fig:trixi_global_overview}
\end{figure*}

Figure~\ref{fig:trixi_global_overview} presents an overview of the basic
components of \trixi. The most important structure is the semidiscretization,
which bundles all information about the spatial part. The mathematical-physical
model is determined by the \lstinline{equations}, the \lstinline{initial_condition},
\lstinline{boundary_conditions}, and possible \lstinline{source_terms}. The
\lstinline{solver} describes purely numerical parameters determining the specific
discretization method such as discontinuous Galerkin or finite difference
methods, kinetic energy preserving or shock capturing approaches. Finally, the
\lstinline{mesh} has a necessarily hybrid role including information about the
spatial domain and its discretization.

Adding information about the time span $[0, T]$, a semidiscretization of \trixi
can be converted into an ODE problem, which can be solved by methods from
OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
allows us to provide extended functionality for \trixi without modifying any
main loop. In particular, diverse tasks such as input/output operations, adaptive
mesh refinement (AMR), and positivity preserving limiting are implemented using
callbacks.

\subsection{Main features of \trixi}

@@

Finally, the different mesh types available in \trixi allow further optimizations.
For example, if the computational domain is essentially a cube (or square, line)
without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
be used. Depending on the relative computational effort of the discretization
(such as the polynomial degree and the choice of the volume terms), the
Cartesian \lstinline{TreeMesh} can be \ca \SI{10}{\percent}--\SI{25}{\percent}
more efficient than the curvilinear meshes in \trixi.

@@

TODO %TODO
\begin{itemize}
  \item Packages can really be used together at low (or no) cost
  \item Reproducibility
  \item Good software development practices (CI, tests, version management)
  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
  \item more nice features
  \item Can be extended: Multi-physics with 350 lines of code
  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
\end{itemize}
